---
title: "rmaADM"

output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")
library(knitr)
library(dplyr)
library(devtools)
load_all()

```

# rmaADM

The `rmaADM` package provides easy access to USDA Risk Management Agency (RMA) Actuarial Data Master (ADM) datasets. This package downloads and provides cleaned, processed versions of RMA actuarial data for crop insurance analysis across multiple years (2011-2025).

## Installation

You can install the development version of rmaADM from GitHub:

```{r install, eval=FALSE}
# Install from GitHub
remotes::install_github("dylan-turner25/rmaADM")
```

## Usage

The primary exported function of rmaADM package is `get_adm_data`. This function retrieves precompiled data sets that are stored in the package's GitHub repository as a release. Below is an example of how to use this function:
```{r example, echo=TRUE, eval = FALSE}

# install the rmaADM package from GitHub
remotes::install_github("dylan-turner25/rmaADM", force = T)

# load the rmaADM package
library(rmaADM)

# load the 2012 base rate data set. If this is the first 
# time downloading, you'll see a message about the dataset 
# being downloade from github.
df <- get_adm_data(year = 2012, dataset = "baserate")

# the data is cached in the R library folder for the rmaADM package
# if you run the same command again, it will not download the data again
# and will load much faster.
df <- get_adm_data(year = 2012, dataset = "baserate")

# if you want to clear the cache and force the data to be downloaded again,
# you can use the clear_rmaADM_cache function.
clear_rmaADM_cache()

# running the same command again will now download the data again
df <- get_adm_data(year = 2012, dataset = "baserate")


# the get_adm_data function is case-insensitive for dataset names
# and can use the name or record code. All of the following return the 
# same data for the year 2012.
get_adm_data(year = 2012, dataset = "BaseRate")
get_adm_data(year = 2012, dataset = "base_RaTe")
get_adm_data(year = 2012, dataset = "A01010")
get_adm_data(year = 2012, dataset = "a01010")


# to see all the available datasets you can run the helper function list_data_assets
all_data <- rmaADM:::list_data_assets()
```

## Available Datasets

The package provides access to multiple ADM datasets across years 2011-2025:

- **BaseRate (A01010)**: Base rates for crop insurance
- **Price (A00810)**: Commodity prices  
- **SubsidyPercent (A00070)**: Premium subsidy percentages
- **InsuranceOffer (A00030)**: Insurance offer information
- **Beta (A01020)**: Beta values for rate calculations
- **ComboRevenueFactor (A01030)**: Combined revenue factors
- **CoverageLevelDifferential (A01040)**: Coverage level differentials
- **HistoricalRevenueCapping (A01110)**: Historical revenue capping data
- **HistoricalYieldTrend (A01115)**: Historical yield trend data (2016+)
- **AreaCoverageLevel (A01130)**: Area coverage levels (2017+)
- **AreaRate (A01135)**: Area rates (2017+)
- **Date (A00200)**: Date information

```{r datasets, eval=FALSE}
# Example: Get price data for 2024
price_data <- get_adm_data(year = 2024, dataset = "price")

# Example: Get subsidy percentage data for 2023  
subsidy_data <- get_adm_data(year = 2023, dataset = "subsidypercent")

# Example: Get historical yield trends for 2020
yield_trends <- get_adm_data(year = 2020, dataset = "historicalyieldtrend")
```

## Key Functions

- **`get_adm_data(year, dataset)`**: Main function to retrieve ADM data
- **`clear_rmaADM_cache()`**: Clear cached data to force re-download
- **`list_data_assets()`**: List all available data assets
- **`build_county_yield_history()`**: Build county yield history datasets

## Data Coverage

The package covers years 2011-2025, with some datasets only available for certain years:
- Core datasets (BaseRate, Price, SubsidyPercent, etc.): 2011-2025
- HistoricalYieldTrend: 2016-2025  
- Area-level datasets (AreaCoverageLevel, AreaRate): 2017-2025

## License

MIT License

## Author

Dylan Turner (dylan.turner3@outlook.com)

